#Reproducible Research: Peer Assessment 1


```{r}
echo = TRUE
```

## Loading libraries 

```{r}
library(plyr)
library(lattice)
library(ggplot2)
```

## Loading and preprocessing the data

```{r}
unzip("activity.zip")
rdata <- read.csv("activity.csv")
rdata$date <- as.Date(rdata$date)
```

## What is mean total number of steps taken per day?

Plot the histogram of total number of steps taken per day:
```{r}
perday.steps <- tapply(rdata$steps, rdata$date, FUN=sum, na.rm=TRUE)
qplot(perday.steps, binwidth=1200, xlab = "Total Number of Steps Taken Per Day")
```

The mean is:
```{r}
mean(perday.steps, na.rm=TRUE)
```
The median is:
```{r}
median(perday.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?

Plot a time series:
```{r}
ndata <- na.omit(rdata)
raverage <- ddply(ndata, ~interval, summarize, mean=mean(steps))
qplot(x=interval, y=mean, data=raverage, geom="line", xlab="5-Minute Interval", ylab="Number of Steps")
```

The 5-minute interval contains the maximum number of steps on average across all the days:
```{r}
raverage[which.max(raverage$mean), ]
```

## Imputing missing values

How many missing values:
```{r}
sum(is.na(rdata))
```
Make a new histogram:
```{r}
raverage <- aggregate(x = list(steps = rdata$steps), by = list(interval=rdata$interval), FUN="mean", na.rm=TRUE)
names(raverage)[2] <- "meanstep"
nndata <- rdata
for (i in 1:nrow(nndata)) {
  if (is.na(nndata$steps[i])) {
    nndata$steps[i] <- raverage[which(nndata$interval[i] == raverage$interval), ]$meanstep
  }
}
perday.steps <- tapply(nndata$steps, nndata$date, FUN=sum)
qplot(perday.steps, binwidth=1200, xlab = "Total Number of Steps Taken Per Day")
```

Now the mean is:
```{r}
mean(perday.steps)
```
Now the median is:
```{r}
median(perday.steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a factor variable:
```{r}
nndata$weektime <- factor(format(nndata$date, "%A"))
levels(nndata$weektime) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), weekend = c("Saturday", "Sunday"))
```

Make plots for weekdays and weekends:
```{r}
raverages <- aggregate(nndata$steps, list(interval=as.numeric(as.character(nndata$interval)), weektime = nndata$weektime), FUN="mean")
names(raverages)[3] <- "meanstep"
xyplot(raverages$meanstep ~ raverages$interval | raverages$weektime,
       type = 'l', layout = c(1,2),
       xlab = "5-Minute Interval",
       ylab = "Number of Steps")
```

### Conclusion
* Yes, the plot reveals people's activity pattern in weekdays is different to that in weekends.
